Exploring Diamonds Dataset Part 2
=============================================
##First look at data.
```{r}
library(ggplot2)
library(hexbin)
data(diamonds)
str(diamonds)
```
##Scatterplot of Price vs X (length in mm)
```{r}
ggplot(data=diamonds, aes(x = x, y = price)) + geom_point() + 
  xlab("Length in mm") + ylab("Price") + ggtitle("Price vs Length in mm")
```

Same scatterplot without outliers.
```{r}
ggplot(data=diamonds, aes(x = x, y = price)) + geom_jitter(alpha=1/15, position = position_jitter(h=0)) + 
  xlab("Length in mm") + ylab("Price") + ggtitle("Price vs Length in mm") +
  coord_cartesian(xlim=quantile(diamonds$x, probs=c(.025, .975)), ylim=quantile(diamonds$price, probs=c(.025,.975))) 
```

What are observation about the scatterplot of price vs x?

1. It's looks like there are exists quadratic dependency between variables.

2. From plot we can see few groups of diamonds price's with respect to length.

3. There exists outliers.

##Correlations

What is the correlation between price and x ?
```{r}
with(diamonds, cor.test(price, x, method='pearson'))
```

What is the correlation between price and y?
```{r}
with(diamonds, cor.test(price, y, method='pearson'))
```

What is the correlation between price and z?
```{r}
with(diamonds, cor.test(price, z, method='pearson'))
```

##Price vs Depth

```{r}
ggplot(data=diamonds, aes(x=depth, y = price)) + xlab("Depth") +
  ylab("Price") + ggtitle("Price vs Depth") +
  stat_binhex() +
  scale_x_continuous(breaks = seq(min(diamonds$depth), max(diamonds$depth), 2))
```

Most diamonds are between this values of depth.
```{r}
with(diamonds, quantile(depth, probs=c(0.05,0.95)))
```

What's correlation between of depth vs price?
```{r}
with(diamonds, cor.test(depth, price, method='pearson'))
```

For prediction depth isn't so important looking at correlation coefficient, but we can't say that unimportant at all using significant level 5%. P-value is less than .05, so wee can reject hypothesys that true correlation is equal to 0. 

##Price vs Carat

Very intelligent scatterplot.
```{r}
ggplot(data=diamonds, aes(x=carat, y = price)) + xlab("Carat") +
  ylab("Price") + ggtitle("Price vs Carat") +
  stat_density2d(geom="tile", aes(fill=..density..^0.25, alpha=1), contour=FALSE) + 
  geom_point(size=0.5) +
  stat_density2d(geom="tile", aes(fill=..density..^0.25,     alpha=ifelse(..density..^0.25<0.4,0,1)), contour=FALSE) + 
  scale_fill_gradientn(colours = colorRampPalette(c("white", blues9))(256))
```

##Price vs Volume

```{r}
diamonds$volume = with(diamonds, x*y*z)
ggplot(data=diamonds, aes(x=volume, y=price)) + 
  xlab("Volume") + ylab("Price") + ggtitle("Price vs Volume") + 
  geom_jitter(alpha=1/20, position=position_jitter(h=0), color='blue') + 
  coord_cartesian(xlim = c(0, quantile(diamonds$volume, 0.99)))
```

Correlation between Price and Volume.

```{r}
with(diamonds, cor.test(price, volume, method='pearson'))
```

How many diamonds have 0 Volume?
```{r}
table(diamonds$volume == 0)[2]
```

Correlation between Price and Volume without outliers.
```{r}
with(subset(diamonds, volume > 0 & volume < 800), cor.test(volume, price, method='pearson'))
```

##Price vs Volume Smooth
```{r}
ggplot(data=diamonds, aes(x=volume, y=price)) + 
  xlab("Volume") + ylab("Price") + ggtitle("Price vs Volume") + 
  geom_jitter(alpha=1/10, position=position_jitter(h=0)) + 
  coord_cartesian(xlim = c(0, quantile(diamonds$volume, 0.99)), ylim = c(0, max(diamonds$price))) +
  stat_smooth(method = "gam", formula = y ~ s(x), size = 1)
```

##Price vs Volume by Color
```{r}
ggplot(data=diamonds, aes(x=volume, y=price, color = color)) + 
  xlab("Volume") + ylab("Price") + ggtitle("Price vs Volume") + 
  geom_jitter(alpha=1/4, position=position_jitter(h=0)) + 
  coord_cartesian(xlim = c(0, quantile(diamonds$volume, 0.99)), ylim = c(0, max(diamonds$price)))
```

##Mean price by Clarity using dplyr

```{r}
library(dplyr)
diamondsByClarity <- diamonds %>% group_by(clarity) %>%
  summarise(mean_price = mean(price),
            median_price = median(as.numeric(price)),
            max_price = max(price),
            min_price = min(price),
            n = n()) %>%
  arrange(clarity)
diamondsByClarity
```

##Bar chart of mean price

```{r, fig.width=10, fig.height=7}
library(gridExtra)
mean_by_clarity <- diamonds %>% group_by(clarity) %>% summarise(mean_price = mean(price))
mean_by_color <- diamonds %>% group_by(color) %>% summarise(mean_price=mean(price))
mean_by_cut <- diamonds %>% group_by(cut) %>% summarise(mean_price = mean(price))

g1 <- ggplot(data=mean_by_clarity, aes(y=mean_price, x = clarity, fill=clarity)) + geom_bar(stat='identity', colour="black") + xlab("Clarity") + ylab("Mean Price") + ggtitle("Mean price by clarity")
g2 <- ggplot(data=mean_by_color, aes(y=mean_price, x = color, fill = color)) + geom_bar(stat='identity',colour="black") + xlab("Color") + ylab("Mean Price") + ggtitle("Mean price by color")
g3 <- ggplot(data=mean_by_cut, aes(y=mean_price, x = cut, fill = cut)) + geom_bar(stat='identity',colour="black") + xlab("Cut") + ylab("Mean Price") + ggtitle("Mean price by cur")

grid.arrange(g1,g2,g3)
```

##Session Info
```{r} 
sessionInfo()
```

