Exploring Diamonds Dataset Part 3
===================================
##Diamonds Dataset
```{r}
library(ggplot2)
data(diamonds)
str(diamonds)
```

##Price Histograms with Facet and Color

```{r}
f <- ggplot(data=diamonds, aes(x=price, color = cut)) +
  geom_histogram(binwidth=0.1) +
  facet_wrap(~color) + scale_x_log10() +
  xlab("Price") + ylab("Frequency") + ggtitle("Price with Facet and Color")
ggsave("0.jpg", f)
```

##Price vs Table Colored by Cut

```{r}
f <- ggplot(data=diamonds, aes(x=table, y=price, color = cut)) + 
  geom_point() +
  xlab("Table") + ylab("Price") + ggtitle("Price vs Table by Cut") + 
  scale_color_brewer(type = 'qual')
ggsave("1.jpg", f)
```

What is the typical table range for the majority of diamonds of ideal cut?
```{r}
with(subset(diamonds, cut == "Ideal") , quantile(table, probs= c(0.1,0.9)))
```

What is the typical table range for the majority of diamonds of premium cut?
```{r}
with(subset(diamonds, cut=='Premium'), quantile(table, probs=c(0.1,0.9)))
```

##Price vs Volume and Diamond Clarity

```{r}
diamonds$volume <- with(diamonds, x*y*z)
f <- ggplot(data=subset(diamonds,volume < quantile(volume, probs=0.99) & volume > 0), aes(x =volume, y=price, color=clarity)) +
  geom_point() + scale_color_brewer(type = 'div') + scale_y_log10() +
  xlab("Volume") + ylab("Price") + ggtitle("Price vs Volume by Clarity")
ggsave("2.jpg", f)
```

##Price/Carat Binned, Faceted, & Colored

```{r}
f <- ggplot( data = diamonds,aes( x = cut, y = price/carat, color = color)) + 
  geom_point(position = position_jitter(width = 0.4)) +
  facet_wrap(~clarity) + 
  scale_color_brewer(type='div') +
  xlab("Cut") + ylab("Price/Carat")
ggsave('3.jpg', f)
```


##Price vs Carat and Clarity with Transformations

```{r warning=FALSE}
library(scales)
cuberoot_trans <- function() trans_new(
  'cuberoot', transform=function(x) x^(1/3),
  inverse=function(x) x^3)

f <- ggplot(aes(x = carat, y = price, color = clarity), data = diamonds) + 
  geom_point(alpha = 0.5, size = 1, position = 'jitter') +
  scale_color_brewer(type = 'div',
                     guide = guide_legend(title = 'Clarity', reverse = T,
                                          override.aes = list(alpha = 1, size = 2))) +  
  scale_x_continuous(trans = cuberoot_trans(), limits = c(0.2, 3),
                     breaks = c(0.2, 0.5, 1, 2, 3)) + 
  scale_y_continuous(trans = log10_trans(), limits = c(350, 15000),
                     breaks = c(350, 1000, 5000, 10000, 15000)) +
  ggtitle('Price (log10) by Cube-Root of Carat and Clarity')
ggsave('4.jpg', f)

```


##Price vs Carat and Cut with Transformation
```{r warning=FALSE}
f <- ggplot(aes(x = carat, y = price, color = cut), data = diamonds) + 
  geom_point(alpha = 0.5, size = 1, position = 'jitter') +
  scale_color_brewer(type = 'div',
                     guide = guide_legend(title = 'Cut', reverse = T,
                                          override.aes = list(alpha = 1, size = 2))) +  
  scale_x_continuous(trans = cuberoot_trans(), limits = c(0.2, 3),
                     breaks = c(0.2, 0.5, 1, 2, 3)) + 
  scale_y_continuous(trans = log10_trans(), limits = c(350, 15000),
                     breaks = c(350, 1000, 5000, 10000, 15000)) +
  ggtitle('Price (log10) by Cube-Root of Carat and Cut')
ggsave('5.jpg', f)
```


##Price vs Carat and Color with Transformation
```{r warning=FALSE}
f <- ggplot(aes(x = carat, y = price, color = color), data = diamonds) + 
  geom_point(alpha = 0.5, size = 1, position = 'jitter') +
  scale_color_brewer(type = 'div',
                     guide = guide_legend(title = 'Color', reverse = F,
                                          override.aes = list(alpha = 1, size = 2))) +  
  scale_x_continuous(trans = cuberoot_trans(), limits = c(0.2, 3),
                     breaks = c(0.2, 0.5, 1, 2, 3)) + 
  scale_y_continuous(trans = log10_trans(), limits = c(350, 15000),
                     breaks = c(350, 1000, 5000, 10000, 15000)) +
  ggtitle('Price (log10) by Cube-Root of Carat and Color')
ggsave('6.jpg', f)
```


##Build Linear Model for Price

```{r warning=FALSE}
library(memisc)
m1 <- lm(I(log(price)) ~ I(carat^(1/3)), data=diamonds)
m2 <- update(m1, ~ . + carat)
m3 <- update(m2, ~ . + cut)
m4 <- update(m3, ~ . + color)
m5 <- update(m4, ~ . + clarity)
mtable(m1,m2,m3,m4,m5)
```

##Session Info

```{r}
sessionInfo()
```

