Exploring Diamonds dataset
=====================================
##Diamonds
```{r}
library(ggplot2)
data(diamonds)
```
How many observations are in data set?
```{r}
dim(diamonds)[1]
```
How many variables are in data set?
```{r}
dim(diamonds)[2]
```
How many ordered factors are in data set?
```{r}
str(diamonds)
```
What letter represents the best color for a diamond?
```{r}
head(levels(diamonds$color),1)
```

##Price Histogram
```{r}
ggplot(aes(x=price), data=diamonds) + geom_histogram(color='blue', fill='#099DD9', binwidth=100) +
  xlab("Price") + ylab("Frequency") + ggtitle("Histogram of diamonds prices")
```

##Price Histogram Summary
```{r}
summary(diamonds$price)
```

##Diamond Counts
How many diamonds cost less than $500?
```{r}
dim(subset(diamonds, price < 500))[1]
```
How many diamonds cost less than $250?
```{r}
dim(subset(diamonds, price < 250))[1]
```
How many diamonds cost $15,000 or more?
```{r}
dim(subset(diamonds, price >= 15000))[1]
```

##Cheaper Diamonds
```{r}
ggplot(aes(x=price), data=subset(diamonds, price < 1500)) +
  geom_histogram(color='blue', fill='#099DD9', binwidth = 1) +
  xlab("Price") + ylab("Frequency") + ggtitle("Histogram of cheaper diamonds")
```

##Price by Cut Histograms
```{r}
ggplot(aes(x=price), data=diamonds) +
  geom_histogram(color = 'blue', fill='#099DD9', binwidth=1) +
  facet_wrap(~cut, ncol=2, scales="free") + xlab("Price") + ylab("Frequency") +
  ggtitle("Histogram of diamonds by cut")
```

##Price by Cut
```{r}
by(diamonds$price, diamonds$cut, summary)
```

##Price per Carat by Cut
```{r}
ggplot(aes(x=price/carat), data=diamonds) +
  geom_histogram(color='blue', fill='#099DD9', binwidth=.05) +
  facet_wrap(~cut, ncol=2, scales='free') + xlab("Price per Carat") + ylab("Frequency") +
  ggtitle("Price per Carat by Cut") + scale_x_log10()
```

##Price Box Plots
```{r}
ggplot(aes(y=price, x=cut), data=diamonds)+
  geom_boxplot() + xlab("Cut") + ylab("Price") + ggtitle("Boxplot. Price by Cut")

ggplot(aes(y=price, x=clarity), data=diamonds)+
  geom_boxplot() + xlab("Clarity") + ylab("Price") + ggtitle("Boxplot. Price by Clarity")

ggplot(aes(y=price, x=color), data=diamonds)+
  geom_boxplot() + xlab("Color") + ylab("Price") + ggtitle("Boxplot. Price by Color")

```

##Interquartile Range
What is the price range for the middle 50% of diamonds with color D?
```{r}
summary(subset(diamonds, color == "D")$price)
#alternative
quantile(subset(diamonds, color == "D")$price, probs=c(0.25, 0.75))
```
What is the price range for the middle 50% of diamonds with color J?
```{r}
quantile(subset(diamonds, color == "J")$price, probs=c(0.25, 0.75))
```
What is the IQR for diamonds with the best color?
```{r}
IQR(subset(diamonds, color == head(levels(diamonds$color), 1))$price)
```
What is the IQR for diamonds with the worst color?
```{r}
IQR(subset(diamonds, color == tail(levels(diamonds$color), 1))$price)
```

##Price per Carat Box Plots by Color
```{r}
ggplot(aes(y=price/carat, x=color), data=diamonds) +
  geom_boxplot(colour = I("#3366FF")) + xlab("Color") + ylab("Price per Carat") +
  ggtitle("Price per Carat by Color")
```

##Carat Frequency Polygon
```{r}
ggplot(aes(x = carat), data = diamonds) +
  geom_freqpoly(binwidth=0.01, color='blue') +
  xlab("Carat") + ylab("Frequency") + ggtitle("Carat Frequency Polygon")
```

What cara size has frequency greater than 2000?
```{r}
t <- table(diamonds$carat)
t[t>2000]
```

##Session Info
```{r}
sessionInfo()

```


